---
title: "p8105_hw2_hy2580"
author: "haokun yuan"
date: "10/1/2018"
output: github_document
---

```{r setup, include=TRUE}
library(tidyverse)
```

##Problem 1

Import and Clean the Data:
```{r}
# import data and format data with appropriate variable names
nyc_transit_data = read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>% 
# select wanted variables
  select(line, station_name, route1, route2, route3, route4, route5, route6, route7, route8, route9, route10, route11, station_latitude, station_longitude, entry, vending, entrance_type, ada) %>% 
# convert the class of variable "entry" to logical
  mutate(entry = recode(entry,`YES` = TRUE, `NO` = FALSE )) 
```

The nyc_transit_data dataset contains 19 variables, which are line, station, route1, route2, route3, route4, route5, route6, route7, route8, route9, route10, route11, station_latitude, station_longitude, entry, vending, entrance_type, ada


I first import data and use janitor:clean_names to convert the names of the variable, then selected the variable of interest from the original dataset and recode the value of "entry" variable to logical variable. 

There are `r ncol(nyc_transit_data)` columns and `r nrow(nyc_transit_data)` rows in the dataset.

The dataset is not tidy although all the column are variables, rows are observations and every value has a cell, route number has spread across 11 columns.

There are `r count(distinct(nyc_transit_data, station_name, line))` distinct stations.

`r sum(distinct(nyc_transit_data, station_name, line, ada)$ada, na.rm = TRUE)` stations are ADA compliant.

`r filter(nyc_transit_data, vending == "NO", entry =="TRUE") %>% nrow / filter(nyc_transit_data, vending == "NO") %>% nrow` of station entrances/exits without vending allow entrance.

Tidy Data
```{r}
# tidy nyc_transit_data
# make route number and route name distinct variables
nyc_transit_tidy = gather(nyc_transit_data, key = "route_number", value = "route_name", route1:route11)
```

Route A Data
```{r}
#exclude all the route name except A 
nyc_transit_a = filter(nyc_transit_tidy, route_name == "A")
```

There are `r count(distinct(nyc_transit_a,station_name,line))` distinct stations serve the A train.
 
Of the stations that serves A train, `r sum(distinct(nyc_transit_a,station_name,line,ada)$ada, na.rm = TRUE)` of them are ADA compliant.

##Problem 2

Import and Clean Mr Trash Wheel
```{r}
library(readxl)
# import data from excel with specify sheet number = 1, omit column with notes
mr_trash_wheel = read_excel("data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = 1, range = cell_cols("A:N")) %>%
#clean name
  janitor::clean_names() %>% 
# omit rows do not include dumpster-specific data
  drop_na(dumpster) %>% 
# round the number of sports balls.
# convert the class of sports balls to integer
  mutate(sports_balls = as.integer(round(sports_balls, digits = 0)))
```

Import and Restructure Precitation Data
```{r}
# import precipitation data on 2017
prcp_2017 = read_excel("data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = 4, range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  filter(!is.na(total)) %>% 
  mutate(year = 2017)
# import precipitation data on 2016
prcp_2016 = read_excel("data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = 5, range = "A2:B14") %>% 
  janitor::clean_names() %>%
  filter(!is.na(total)) %>% 
  mutate(year = 2016)
# bind prcp_2017 and prcp_2016
prcp = bind_rows(prcp_2017, prcp_2016) %>%
  mutate(month = month.name[month])
```
Mr Trash Wheel contains 285 observations of 14 variables, and the key variables in this dataframe is the detail of the trash each dumpster collected, such as sports_balls, weight_tons. The joint precipitation data of 2017 and 2016 contains 24 observations of 3 variables and the key variables are month, year and total.

The total precipitation in 2017 is `r sum(prcp_2017$total)`

The median number of sports balls in a dumpster in 2016 is `r median(filter(mr_trash_wheel, year == 2016)$sports_balls)`.

##Problem 3

```{r}
# install.packages("devtools")
devtools::install_github("p8105/p8105.datasets")
library(p8105.datasets)
```

Import and Tidy Data
```{r}
brfss_tidy =
  brfss_smart2010 %>% 
  janitor::clean_names() %>%
  select(-class,-topic, -question, -sample_size,-(confidence_limit_low:geo_location)) %>%  
  filter(response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) %>% 
  spread(key = response, value = data_value) %>% 
  janitor::clean_names() %>% 
  mutate(excellent_verygood = (excellent + very_good)/(excellent + very_good + good + fair + poor))   
  #variable unit
  

```


```{r}
state_freq = count(brfss_tidy,locationabbr) %>%
  arrange(desc(n))
```

There are `r count(distinct(brfss_tidy,locationdesc))` unique locations, and those locations are from `r count(distinct(brfss_tidy,locationabbr))` different state. Since the United States has 50 states in addition to Washington DC, we conclude that all the states were represented.

The most observed city is `r state_freq[1,1]`.


In 2002, the median of "Excellent" response value is `r filter(brfss_tidy, year == 2002)$excellent %>% median(na.rm = TRUE)`.


Histogram of "Excellent" value in 2002
```{r}
brfss_2002 = brfss_tidy %>%
  filter(year == 2002)
ggplot(brfss_2002, aes(x = excellent)) +
  geom_histogram()
```

Scatterplot of Excellent proportion NY and Queens(2002-2010)
```{r}
brfss_excellent_ny_queens = brfss_tidy %>% 
    filter(locationdesc %in% c("NY - New York County", "NY - Queens County", year == 2002:2010 ))
ggplot(brfss_excellent_ny_queens, aes(x = year, y = excellent, color = locationdesc)) +
  geom_point()
```



